Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$: limitations imposed by cosmic variance

### 2013-07-31 paper/ms.tex
abstract + intro

\newcommand{\kms}{\,km~s$^{-1}$}
\def\squig{\sim\!\!}
\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}   
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\Lsun}{\mbox{\,$L_{\odot}$}}
\newcommand{\like}{\mathscr{L}}
\newcommand{\bftheta}{\mathbf{\Theta}}
\newcommand{\degree}{\ensuremath{^\circ}}
\def\spose#1{\hbox to 0pt{#1\hss}}
\def\simlt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13C$}}}
\def\simgt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13E$}}}
\font\smcap=cmcsc10

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a
unique and narrow set of values for the parameters in the model. In
particular, it is always possible to find a consistent model with any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq
1.0$.  Nevertheless, in the models we can define three
different families of parameters based on their mass range $\Delta M
\equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ and occupation
fraction $f_{\rm occ}$. The dominant family is composed by models with
a narrow mass range $\Delta M<1.0$ dex, a low occupation fraction
$f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm  max}<
10^{12}$\hMsun.  The existence of this dominant family gives support
to the idea that the most massive dark matter halos at that epoch do
not host the brightest LAEs and that only a small fraction of star
forming galaxies can be actually detected as LAEs.

@@

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter (DM), whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980}

Most models of galaxy formation find that the mass of the halo can be
used to predict properties of the galaxy such as its stellar mass and
star formation rate \citep{Behroozi2012}. Processes that regulate the
star formation cycle are also though to be strongly dependent on its
mass.  For the reasons mentioned above finding the typical dark matter
halo mass hosting LAEs represents a significant step forward to
understand the nature of this population in the context of Lambda Cold
Dark Matter ($\Lambda$CDM) paradigm.

Some theoretical approaches to this problem have been based on
ab-initio approach modeling. Starting from the DM halo population, the
corresponding intrinsic star formation properties are inferred
together with other statistics such as the luminosity function, the
correlation function and the equivalent width distributions. Such
modelling has been implemented from analytic considerations,
semi-analytic models  and  full N-body hidrodynamical simulations
\citep{Laursen2007, Dayal2009, ForeroRomero2011, Yajima2012, Orsi2012, Soler2012}.

Added to the uncertainties in the astrophysical processes describing
star formation in galactic populations, a highly debated step in this
approach is the calculation of the fraction of \ly photons
that escape the galaxy to the observer. Given the resonant nature of
the \lya line, the radiative transfer of \ly is sensitive to the
density, temperature, topology and kinematics of the neutral Hydrogen
in the interstellar medium (ISM) \citep{Neufeld1991,ForeroRomero2011,
  Dijkstra2012,Laursen2013,Orsi2012}.

This complexity can be tackeld with monte-carlo simulations for the
radiative transfer a required tool to obtain physically sound results,
although the degeneracy in the physical parameters involved in the
problem makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

Another approach uses the spatial information to infer the typical
mass of the hosting dark matter halo. This uses the fact that in cold
dark matter cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010}. In this
studies the observations were done on single fields of $\sim 1$
deg$^{2}$ and the conclusions derived on the halo host mass from
clustering information do not delve to deeply into the impact of
cosmic variance.

Recently \cite{Yamada2012} observed a wide area a bit more than ten
times the sky area covered so far, all of it with the same instrument
and the same criteria to reduce the observational data and produce LAE
catalogs. This large data-set allows a first observational study on the
expected cosmic variance effects in LAEs at $z=3.1$. This motivate us
to perform such a study from a theoretical persepective and study the
effects of the explicit inclusion of cosmica variance effects by using
a large N-body cosmological simulation. This
allows us to build mock fields resembling the observational geometry
and and directly expore the compatibility of a model with
observations.

In this \documentname, besides studying the impact of cosmic variance,
we implement a method to populate the DM halos with LAEs that bypasses
all the uncertainty involved in the estimation of star formation rates
and \ly escape fraction. Our model only defines whether a DM hosts a
detectable LAE or not. It does not predict a \lya luminosity. By using
this model we want to define the preferred mass range of DM halos
hosting these galaxies and the fraction of such halos that actually
can host a LAE.  Once the mock catalogs are constructed in this model
we proceed with a direct comparison against the statistics derived
from observations.

This \documentname is structured as follows. In the next section we present
the simulation and model used to produce the mock catalogs and the criteria
used to compare them against observations. In \SS \ref{sec:results} we
present the main results in terms of the parameters of our model. We
continue with a discussion of these results under the light of other
observational results for LAEs and abundance matching models for
galaxy formation before presenting our conclusions in \SS
\ref{sec:conclusions}.

@@

At the low mass end, such cut can be readily interpreted in terms of the
minimal halo star formation rate needed to produce the necessary \lya
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increasing with halo
mass, the cut at higher halo masses requires a different
explanation. There are two different physical scenarios to explain
this.

>>>>

\newcommand{\kms}{\,km~s$^{-1}$}
\def\squig{\sim\!\!}
\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}   
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\Lsun}{\mbox{\,$L_{\odot}$}}
\newcommand{\like}{\mathscr{L}}
\newcommand{\bftheta}{\mathbf{\Theta}}
\newcommand{\degree}{\ensuremath{^\circ}}
\def\spose#1{\hbox to 0pt{#1\hss}}
\def\simlt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13C$}}}
\def\simgt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13E$}}}
\font\smcap=cmcsc10

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a unique and narrow set of values
for the parameters in the model. In particular, it is always possible
to find a model that is consistent with observations and has any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq 1.0$
.  Nevertheless, in the models we can define three different families
of parameters based on their mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ and occupation fraction
$f_{\rm occ}$. The dominant family is composed by models with a narrow
mass range $\Delta M<1.0$ dex, a low occupation fraction $f_{\rm
  occ}\leq 0.3$ and a maximum mass $M_{\rm  max}< 10^{12}$\hMsun.  The
existence of this dominant family gives support to the idea that the
most massive dark matter halos at that epoch do not host the brightest
LAEs and that only a small fraction of star forming galaxies can be
actually detected as LAEs.

@@

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter (DM), whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980,SpringelNature05}.

Most models of galaxy formation find that halo mass can be
used to predict galaxy properties such as the stellar mass and
star formation rate \citep{Behroozi2012}. Processes that regulate the
star formation cycle are also though to be strongly dependent on its
mass.  For these reasons, finding the typical dark matter halo mass
hosting LAEs represents a significant step forward to understand the
nature of this galaxy population in the context of Lambda Cold Dark
Matter ($\Lambda$CDM) paradigm.

Some theoretical approaches to this problem have been based on
ab-initio approach. Starting from the DM halo population, the
corresponding intrinsic star formation properties are inferred
together with other statistics such as the luminosity function, the
correlation function and the equivalent width distributions. Such
modelling has been implemented from analytic considerations,
semi-analytic models  and  full N-body hidrodynamical simulations
\citep{Laursen2007, Dayal2009, ForeroRomero2011, Yajima2012, Orsi2012, Soler2012}.

In addition to the uncertainties in the astrophysical processes describing
star formation in galactic populations, is the calculation of the fraction of \ly photons
that escape the galaxy to the observer is another debated step. Given the resonant nature of
the \ly line, the radiative transfer of \ly is sensitive to the
density, temperature, topology and kinematics of the neutral Hydrogen
in the interstellar medium (ISM). \citep{Neufeld1991,ForeroRomero2011,
  Dijkstra2012,Laursen2013,Orsi2012}.

This problem can be tackeld with Monte Carlo simulations for the
radiative transfer, however the complexity of the relevant physical
processes makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

Another approach to infer the typical DM halo mass of halos hosting
LAEs is based on the spatial information. This uses the fact that in cold
dark matter cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010}. In this
studies the observations were done on single fields of $\sim 1$
deg$^{2}$ and the conclusions derived on the halo host mass from
clustering information do not delve to deeply into the impact of
cosmic variance.

Recently \cite{Yamada2012} observed a wide area a bit more than ten
times the sky area covered so far by individual campaigns. The
observations were done with the same instrument and same criteria to
reduce the observational data and produce LAE catalogs. This large
data-set allows us to make the first study on the expected cosmic
variance effects in LAEs at $z=3.1$ and its impact on the
determination of the halo mass and occupation fraction.

In this \documentname to study the impact of cosmic variance,
we implement a method to populate the DM halos in cosmological
simulations with LAEs. This approach bypasses all the uncertainty
involved in the estimation of star formation rates and \ly escape
fraction. The model only considers whether a DM can host a
detectable LAE or not. It does not predict a \ly  luminosity. Once the
mock catalogs are constructed  we proceed with a direct comparison
against the statistics derived from observations. This allows us 
to find the preferred mass range of DM halos hosting these galaxies
and its occupation fraction.

This \documentname is structured as follows. In the next section we present
the simulation and model used to produce the mock catalogs and the criteria
we use to compare them against observations. In \SS \ref{sec:results} we
present the main results for the halo mass and occupation fraction. We
continue with a discussion of these results under the light of other
observational and theoretical results. Finally, we present our
conclusions in \SS
\ref{sec:conclusions}.

@@

At the low mass end, such cut can be readily interpreted in terms of the
minimal halo star formation rate needed to produce the necessary \ly 
luminosity to be above a given detection threshold.  However, under
the reasonable assumption of star formation rate increasing with halo
mass, the cut at higher halo masses requires a different
explanation. There are two different physical scenarios to explain
this.
